Global patterns of forest autotrophic carbon fluxes

### 2020-02-14 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS; *Piao et al. 2010. Forest annual carbon cost: a global‐scale analysis of autotrophic respiration. Ecology 91(3), 652-661* **). Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity ([Waide et al. 2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**). However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.

**FACF varies with latitude**, showing a general trend of decreasing FACF with latitude [@beer_terrestrial_2010; @jung_global_2011]. Studies agree that FACF is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011]. However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011];[Bagdley et al. 2019](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729)). This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007]. {**condense the previous two sentences into one-- both primarily eddy flux**} In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017]. Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].

@@

**FACF can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011]. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity (Liang et al. 2019. Positive biodiversity-productivity relationship predominant in global forests. Science 354, Issue 6309, aaf8957), management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF. On a global scale, we expect that FACF is most strongly influenced by broad climatic gradients. There is evidence that FACF also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways. Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and FACF is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between primary productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on FACF.

>>>>

**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS**) [@piao_forest_2010]. Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**). However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.

**FACF varies with latitude**, showing a general trend of decreasing FACF with latitude [@beer_terrestrial_2010; @jung_global_2011]. Studies agree that FACF is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011]. However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions. Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships. For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019]. In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017]. Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].

@@

**FACF can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011]. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF. On a global scale, we expect that FACF is most strongly influenced by broad climatic gradients. There is evidence that FACF also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways. Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length. Some studies have suggested that the apparently strong relationship between MAT and FACF is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Kerkhoff et al. [-@kerkhoff_plant_2005] and Michaletz et al. [-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between primary productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on FACF.
